### 2022-09-28 chapters/optimization/python-dir/python.tex

To answer these questions, we report on 4 case studies that intend to answer these research questions.
First, we study the energy behavior of Python in two application contexts: machine learning (cf. \Cref{sec:webdev}) and web applications (cf. \Cref{sec:ml}).
Then, we dive deeper into the energy consumption of Python core structures, before concluding with the impact of parallelism on energy consumption.

\subsection{Python for Machine Learning}\label{sec:ml}
Machine learning is becoming an integral part of our daily lives, growing more potent and energy-hungry each year.
As machine learning can have a significant impact on climate change, it is vital to investigate mitigation techniques.

>>>>

To answer these questions, we report on 4 case studies that intend to answer these research questions.
First, we study the energy behavior of Python in two application contexts: machine learning (cf. \Cref{sec:webdev}) and web applications (cf. \Cref{sec:ml}).
Then, we dive deeper into the energy consumption of Python core structures before concluding with the impact of parallelism on energy consumption.

\subsection{Python for Machine Learning}\label{sec:ml}
Machine learning is becoming an integral part of our daily lives, growing more potent and energy-hungry each year.
As machine learning can significantly impact climate change, it is vital to investigate mitigation techniques.

### 2022-09-28 chapters/programming_languages/rpc/rpc.tex

\subsection{Introduction}
With the success of Internet and the emergence of cloud technologies, many communication protocols compete to take the lead.
In particular, most of the software architectures are now based on multi-services and micro-service technologies.
And, to cope with the higher versatility of developers, multiple companies choose to open their services to different programming languages.
Interestingly, this approach intends to take advantage of each programming language to satisfy specific requirements.
However, the challenge nowadays is to make the bridge between those platforms.
We have many initiatives, such as OpenAPI that try to create a taxonomy for RESTful APIs, while other approaches implement all the different interfaces of the protocol by themselves, such as RPC.

@@

\subsubsection{Stress Free Mode}
In this mode, the compiled languages tend to consume fewer resources (average power).
JVM-based languages tend to consume more energy, especially Scala.
However, we do not observe the same behavior when it comes to efficiency.
Unlike the other interpreted programming languages, PHP performances could be compared to the compiled ones, such as CPP or GO, and even better to some others, such as Swift.
JVM-based languages tend to have better performances than the interpreted ones.
Furthermore, OpenJDK has shown more efficiency than GraalVM [].
Overall, we can have 3 groups when it comes the cost of each request:
\begin{itemize}
    \item Energy-efficient class: C++, GO, RUST, ELIXIR, and PHP,
    \item Middle class: Most of the interpreted languages and VM-based ones,
    \item Energy-greedy class: Crystal and Scala.
\end{itemize}

\subsubsection{Escalation Mode}
In this mode, the behavior of the server depends on the payload. We observe three cases:
\begin{enumerate}
    \item Drop in performances without an increased power, such as .Net core, Java micronaut, Crystal, and Dart.
          In this case, the server keeps using the same resources, and sometimes less, because it takes more time to handle the fewer requests.
          This class of languages tends to be the most energy-consuming when it comes to the cost per request;
    \item Increase in power without affecting the performances, such as Go or .Net.
          The energy consumption of a single request, is affected slightly but still increases;
    \item Increase in power and drop in performances.
          Despite the increase of power consumption, the server becomes slightly slower, which increases the energy cost per request.
          This cost is still better than the first case, which concludes that the servers in the first category are on the verge of breaking.
\end{enumerate}

We can mention the case of Elixir that keeps scaling despite the lack of performances compared to other compiled languages (Go, CPP).

\subsubsection{Broken state mode }
Only four of the 25 configurations could parse the 10 MB files, and only 1 from those could achieve a 76\% acceptance rate which is Elixir, the other 3 had less than 3\% success rate (Rust, Swift and Dart).
The rest could be divided into two categories:
\begin{itemize}
    \item \textsf{Timeout} where requests took too much time that the client canceled them, in this category we find most of dynamic codes, such as OpenJDK and Kotlin,
    \item \textsf{Size of request} exceeded the maximum size when the implementation could not handle requests with large size, as observed with .Net, Go, .Net core, CPP, PHP, Scala, Nodejs, Ruby, Python.
\end{itemize}

@@

In each programming category, we observe both energy-efficient and energy-greedy behaviours.
Therefore, we conclude that it depends more on the implementation of the library itself, rather than the category of the programming language.
Scala and Kotlin are an excellent example to support this hypothesis, as both of them run on the same virtual machine as Java (OpenJDK 16.1).
Yet, their average power is 130\% higher than the Java implementation.

\paragraph{Stressed mode}
Although the same classes remained the same, not all the languages had the same evolution and here we can clearly observe a correlation with the category of the programming language rather than the implementation itself.
We can clearly highlight that VM-based languages have a significant increase (double) in the average power consumption after they receive more than 100 concurrent clients.
Except PHP, all the interpreted and compiled languages preserved their energetic behaviour.
Our hypothesis points to the JIT, since it compiles the code and makes it run faster, hence stressing the CPU.
A interesting behaviour has been noticed for the GraalVM: the decrease of energy consumption when increasing the number of the clients.
This is related to the drop of the performances, which was probably due to the bottleneck situation where the GraalVM could not handle more than 100 clients simultaneously.

\subsubsection{Performance Behaviour}
In this section, we study only the number of requests per seconds processed by the server without looking at its energy.
We consider three observable variables:
\begin{itemize}
    \item \textsf{Satisfaction ratio}: how many requests have been satisfied among the total requests,
    \item \textsf{Request Per Seconds}: The number of the requests that have been answered from the server,
    \item \textsf{Tail Latency at 99\%}: one of the best metrics to evaluate the performances of a server.
\end{itemize}

\paragraph{Satisfaction ratio}
Most of the considered frameworks satisfy all the requests, by either reducing the number of requests per second or by increasing the processing time.
However, there are some frameworks that have chosen a different approach, such as Dart or Scala, where the choice was to keep a certain limit of latency even if not all the requests are answered.
Furthermore, we tend to observe this behaviour among other frameworks, such as Python or Asynchronous NodeJs, when the number of the client exceeds 800.

@@

\begin{figure}[!hbt]
    \begin{center}
        \includegraphics[width=1.2\linewidth]{imgs/rps_payload}
    \end{center}
    \caption{Number of requests per second based on the request size}\label{fig:rps_payload}
\end{figure}
Most of the servers hit their RPS limit after 5 clients and 100 clients for vm based servers, and after this the number keeps constant, which will decrease the average RPS per client.
.Net server is the most performant, followed by Java and Go, while Python and Ruby are the least performant.

\paragraph{Tail Latency}
The increase in the number of requests per second, does not necessarily mean a lower latency.
As one can notice in Fgirue~\ref{fig:tail99_clients}, until the 1000 clients, Go provides the least latency beside .Net.
GraalVM provides the highest latency, on average.
However, Dart tends to become slower when we increase the number of clients, until we pass the 600 simultaneous clients, and there it changes its behaviour, instead of satisfying most the requests it notify the clients directly that the server is saturated, hence a drop in satisfaction ratio, and an amelioration for the average latency.

@@

\subsubsection{Energy Per Request}
Now, after we made a separation between the energy and the performances, we have seen that most of the performance servers tend to be energy-greedy, so we propose to investigate this trade-off between energy and performances.
To do so, we report on an average cost of a single request, in Joules.
Except GraalVM when the cost of the a single request increases with when we add more clients, all the frameworks report on a constant cost, Java, .net and go are the most energy efficient, while Python and Ruby may cost up to 10x more.
Therefore, we conclude that the number of clients does not impact the energy significantly.
Then, we study how the payload size of the requests impact the energy consumption of the framework.

>>>>

\subsection{Introduction}
With the success of the Internet and the emergence of cloud technologies, many communication protocols compete to take the lead.
In particular, most software architectures are now based on multi-services and micro-service technologies.
And, to cope with the increased versatility of developers, multiple companies choose to open their services to different programming languages.
This approach is interesting because it wants to use the best parts of each programming language to meet different needs.
However, the challenge nowadays is to make the bridge between those platforms.
We have many initiatives, such as OpenAPI, that try to create a taxonomy for RESTful APIs, while other approaches implement all the different interfaces of the protocol by themselves, such as RPC.

@@

\subsubsection{Stress Free Mode}
In this mode, the compiled languages tend to consume fewer resources (average power).
JVM-based languages tend to consume more energy, especially Scala.
However, we do not observe the same behavior when it comes to efficiency.
Unlike the other interpreted programming languages, PHP's performance can be compared to the compiled ones, such as CPP or GO, and even better to some others, such as Swift.
JVM-based languages tend to have better performance than interpreted ones.
Furthermore, OpenJDK has shown more efficiency than GraalVM.
Overall, we can have 3 groups when it comes the cost of each request:
\begin{itemize}
    \item Energy-efficient class: C++, GO, RUST, ELIXIR, and PHP,
    \item Middle class: Most of the interpreted languages and VM-based ones,
    \item Energy-greedy class: Crystal and Scala.
\end{itemize}

\subsubsection{Escalation Mode}
In this mode, the behavior of the server depends on the payload. We observe three cases:
\begin{enumerate}
    \item drop in performances without an increased power source, such as Net core, Java Micronaut, Crystal, and Dart.
    In this case, the server keeps using the same resources, and sometimes less, because it takes more time to handle the fewer requests.
    This class of languages tends to be the most energy-consuming when it comes to the cost per request;
    \item increase in power without affecting the performance, such as Go or.Net.
    The energy consumption of a single request is affected slightly but still increases.
    \item Increase in power and a drop in performance.
    Despite the increase in power consumption, the server becomes slightly slower, which increases the energy cost per request.
    This cost is still better than the first case, which concludes that the servers in the first category are on the verge of breaking.
\end{enumerate}

We can mention the case of Elixir that keeps scaling despite the lack of performance compared to other compiled languages (Go, CPP)..

\subsubsection{Broken state mode }
Only four of the 25 configurations could parse the 10 MB files, and only one of those could achieve a 76\% acceptance rate, which is Elixir. The other 3 had less than a 3\% success rate (Rust, Swift, and Dart).
The rest could be divided into two categories:
\begin{itemize}
    \item \textsf{Timeout} where requests took too much time that the client canceled them, in this category we find most of dynamic codes, such as OpenJDK and Kotlin,
    \item \textsf{Size of request} exceeded the maximum size when the implementation could not handle requests with large size, as observed with .Net, Go, .Net core, CPP, PHP, Scala, Nodejs, Ruby, Python.
\end{itemize}

@@

In each programming category, we observe both energy-efficient and energy-greedy behaviors.
Therefore, we conclude that it depends more on the implementation of the library itself than on the category of the programming language.
Scala and Kotlin are excellent examples to support this hypothesis, as both of them run on the same virtual machine as Java (OpenJDK 16.1).
Yet, their average power is 130\% higher than the Java implementation.

\paragraph{Stressed mode}
Although the same classes remained the same, not all the languages had the same evolution, and here we can clearly observe a correlation with the category of the programming language rather than the implementation itself.
We can easily point out that the average amount of power used by VM-based languages doubles when they have more than 100 clients at the same time.
Except PHP, all the interpreted and compiled languages preserved their energetic behaviour.
Our hypothesis points to the JIT, since it compiles the code and makes it run faster, hence stressing the CPU.
An interesting behavior has been noticed for the GraalVM: the decrease in energy consumption when increasing the number of clients.
This is related to the performance drop, which was probably due to the bottleneck situation where the GraalVM could not handle more than 100 clients simultaneously.

\subsubsection{Performance Behaviour}
In this section, we study only the number of requests per second processed by the server without looking at its energy.
We consider three observable variables:
\begin{itemize}
    \item \textsf{Satisfaction ratio}: how many requests have been satisfied among the total requests,
    \item \textsf{Request Per Seconds}: The number of the requests that have been answered from the server,
    \item \textsf{Tail Latency at 99\%}: one of the best metrics to evaluate the performances of a server.
\end{itemize}

\paragraph{Satisfaction ratio}
Most of the considered frameworks satisfy all the requests by either reducing the number of requests per second or increasing the processing time.
However, there are some frameworks that have chosen a different approach, such as Dart or Scala, where the choice was to keep a certain limit of latency even if not all the requests are answered.
Furthermore, we tend to observe this behavior among other frameworks, such as Python or Asynchronous NodeJS, when the number of clients exceeds 800.

@@

\begin{figure}[!hbt]
    \begin{center}
        \includegraphics[width=1.2\linewidth]{imgs/rps_payload}
    \end{center}
    \caption{Number of requests per second based on the request size}\label{fig:rps_payload}
\end{figure}
Most of the servers hit their RPS limit after 5 clients and 100 clients for VM-based servers, and after this the number keeps constant, which will decrease the average RPS per client.
. The Net server is the most performant, followed by Java and Go, while Python and Ruby are the least performant.

\paragraph{Tail Latency}
Even if the number of requests per second goes up, that doesn't mean the latency will go down.
As one can notice in figure~\ref{fig:tail99_clients}, until the 1000 clients, Go provides the least latency besides.Net.
GraalVM provides the highest average latency.
However, Dart tends to become slower when we increase the number of clients, until we pass the 600 simultaneous clients, and then it changes its behavior. Instead of satisfying most of the requests, it notifies the clients directly that the server is saturated, resulting in a drop in satisfaction ratio and an amelioration in average latency.

@@

\subsubsection{Energy Per Request}
Now, after we made a separation between the energy and the performances, we have seen that most of the performance servers tend to be energy-greedy, so we propose to investigate this trade-off between energy and performances.
To do so, we report the average cost of a single request in joules.
Except GraalVM when the cost of the a single request increases with when we add more clients, all the frameworks report on a constant cost, Java, .net and go are the most energy efficient, while Python and Ruby may cost up to 10x more.
Therefore, we conclude that the number of clients does not significantly impact the energy consumption.
Then, we study how the payload size of the requests impacts the energy consumption of the framework.
